Macro drop and droplet profile detection during PGMAW process by image processing library

### 2010-12-11 articles/MIG/doc_broullion/latex/paper-drop.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%    INSTITUTE OF PHYSICS PUBLISHING                                   %
%                                                                      %
%   `Preparing an article for publication in an Institute of Physics   %
%    Publishing journal using LaTeX'                                   %
%                                                                      %
%    LaTeX source code `ioplau2e.tex' used to generate `author         %
%    guidelines', the documentation explaining and demonstrating use   %
%    of the Institute of Physics Publishing LaTeX preprint files       %
%    `iopart.cls, iopart12.clo and iopart10.clo'.                      %
%                                                                      %
%    `ioplau2e.tex' itself uses LaTeX with `iopart.cls'                %
%                                                                      %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%
% First we have a character check
%
% ! exclamation mark    " double quote  
% # hash                ` opening quote (grave)
% & ampersand           ' closing quote (acute)
% $ dollar              % percent       
% ( open parenthesis    ) close paren.  
% - hyphen              = equals sign
% | vertical bar        ~ tilde         
% @ at sign             _ underscore
% { open curly brace    } close curly   
% [ open square         ] close square bracket
% + plus sign           ; semi-colon    
% * asterisk            : colon
% < open angle bracket  > close angle   
% , comma               . full stop
% ? question mark       / forward slash 
% \ backslash           ^ circumflex
%
% ABCDEFGHIJKLMNOPQRSTUVWXYZ 
% abcdefghijklmnopqrstuvwxyz 
% 1234567890
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
\documentclass[12pt]{iopart}
\newcommand{\gguide}{{\it Preparing graphics for IOP journals}}
%Uncomment next line if AMS fonts required
%\usepackage{iopams}  
\begin{document}

@@

\address{IOP 
Publishing, Dirac
House, Temple Back, Bristol BS1 6BE, UK}
\ead{custserv@iop.org}
\begin{abstract}
This document describes the  preparation of an article using \LaTeXe\ and 
\verb"iopart.cls" (the IOP \LaTeXe\ preprint class file).
This class file is designed to help 
authors produce preprints in a form suitable for submission to any of the
journals published by IOP Publishing.
Authors submitting to any IOP journal, i.e.\ 
both single- and double-column ones, should follow the guidelines set out here. 
On acceptance, their TeX code will be converted to 
the appropriate format for the journal concerned.

@@

\section{Introduction}
A better comprehension of welding objects behavior presents in a PGMAW process
 as macro drop and metal transfer drop (droplet) could help to improve numerical simulations
 and enhance welding quality in manufacturing process \cite{LIN}, \cite{WU1}.
% An example to this are the problems linked to productivity. To some process,
 For example, the
 increase of productivity means often increase of weld speed,
 however some welding defects related to bad shape, as the hump formations, limit the maximum weld 
 speed \cite{CHAPUIS,Mendez} and therefore the productivity. 
 In a PGMAW process, the weld quality is strongly relates with the metal transfer  stability 
 (the droplet deposition). The control of metal transfer process %implies control of
 will improve weld quality \cite{WANG}. In both example cases a macro drop kinetic
  study and a droplet dynamic analysis during static PGMAW process could help to better
 understand some GMAW related phenomena as humping formation \cite{CHO}, droplet 
  fluctuations \cite{WANG} that guarantee the welding quality. 
 In both case, the kinetic and the dynamic analysis, required the monitoring \cite{} and the 
  geometrical analysis of macro drop and droplets in order to extract 
  qualitative information of these welding objects. 
  There have been many studies on visual sensing techniques for observing weld pool 
  image \cite{BAE} and metal transfer process during welding \cite{LIN}. Optical sensors like high speed 
 cameras and lighting systems have been widely use in GMAW process to realize 
 image acquisition \cite{GUANGJUN},
   control process \cite{BAE}, parametric studies \cite{BALSAMO} and droplet dynamics analysis\cite{LIN}.
   Image processing plays a critical role in 
   extracting useful information from visual scenes \cite{WANG}.
 Nevertheless the strong interference from the arc lightning required more than standard 
   image treatment to analyses the raw images of the welding process \cite{NORDBRUCH}. 
Previous work has shown that is possible to perform geometrical 
   analysis in weld pools or droplets \cite{WU2}.
   Parameters such as macro drop or droplets surface, volumes 
or height has been measure using different and 
   specifics processing images algorithms \cite{WANG}, \cite{WU2}, \cite{SEED},
 \cite{NORDBRUCH}. However, to date, effectively automatic images processing
    of metal transfer has not been developed, possibly due to the  difficulty involved 
for welding researchers \cite{WANG}.

@@

\section{Problems description}
Bradstreet was the first researcher to experimentally study the bead hump formation \cite{CHO}. 
Humping was defined as the series of undulations of the weld bead. Further studies as computer 
simulation made by CHO et al. \cite{CHO} shown a relation between the droplet momentum deposition
 and the hump size. Same relations has been found between metal transfer
 process and welding quality \cite{}. Then the shape and size kinetic 
analysis of macro drop and droplets could help to better understand 
 and enhance the PGMAW process. To monitor the shape and size of these welding 
objects a 2D approach would be sufficient. Therefore a shadowgraphy technique, or back lighting,
  is the natural choice to record the droplets and macro drops profiles \cite{BALSAMO}, furthermore is 
  one of the only way to access these quantities. Due to the arc light 
interference and the relatively high 
  speed of wire feed process, a high speed camera and an effective image processing algorithms are 
  required \cite{WANG}. The typically frame rate acquisition to metal 
transfer drop kinetic analysis is $3000$ 
  per second \cite{WANG}. Therefore the algorithms have to be able 
to extract the geometrical information (area, size and others)
    from a huge amount of data. 
In addition, the voltage and current signals are directly related 
   with the droplet formation at the wire \cite{BALSAMO}. Good 
synchronizations methods have to be applied between shadowgraphies 
   frames and electrical signals \cite{chapuis}.

\section{Image processing background} 
Some definitions and brief description of generals principles used in
 image processing are required to better understand how is work the 
algorithms utilized by the erCv library.

@@

\subsection{Filters}\label{filters}
The filters are algorithms that operate as mathematical functions $f$ above the $X$, $Y$ or both axis of the image ($Z = f(X,Y)$,  
$f(X)$ or $f(Y)$)  modifying his greyness value or $Z$ component. Different kind of filters can be mentioned as median, Gaussian, impulse, adaptive and others. The impulse filters such Canny are 
widely use to edge detection \cite{COCQUEREZ}. These filter have an impulse response to most important greyness gradient in the 
image; this allows the filter a better edges localisation (see figure \ref{photo-explication-filter}). For this reason, Canny filter is widely 
use at erCv library to detect the welding elements edges. However, it is sensible to noise or secondary greyness gradients in the image
 and, in consequence, it have some difficulties to define closer surface.

@@

\begin{figure}[h!]
\begin{center}    
\subfigure[Weld pool image in static GTAW process]{\label{photo-explication-patron}\includegraphics[width=3.5cm,height=3.5cm]{photo-explication-patron.png}}
\subfigure[Canny filter treatment]{\label{photo-explication-filters}\includegraphics[width=3.5cm,height=3.5cm]{photo-explication-filters.png}}\\
\subfigure[Snake treatment]{\label{photo-explication-snake}\includegraphics[width=3.5cm,height=3.5cm]{photo-explication-snake.png}}
\subfigure[Segmentation image by 2 cluster sample comparison]{\label{photo-explication-segmentation}\includegraphics[width=3.5cm,height=3.5cm]{photo-explication-segmentation.png}}
\end{center}
\caption{{\small Samples of different methods technique for image processing}}
\label{photo-explication}
\end{figure}

@@

\begin{figure}
\begin{center}
\includegraphics[width=7cm,height=4cm]{schema-erCv.png}
\caption{{\small Flow diagram of erCv library composition}}
\label{schema-erCv}
\end{center}
\end{figure}
\end{itemize}

@@

\begin{table}
begin{center}
\begin{tabular}{|cc|}
\hline
Welding wire type & ER70S \\ 
Wire diameter ($mm$) or $drw$ & 1 \\  
Contact tip to work distance ($mm$) & 20 \\
Shielding gaz flowrate ($l/min$) & 18 \\ \hline
\end{tabular}
\caption{{\small Variable welding parameters used in experiments}}
\label{table-parameters-change}
\end{center}
\end{table}

@@

\begin{figure}
\begin{center}
\includegraphics[width=7cm,height=4cm]{schema-erCv.png}
\caption{{\small Profile schematics of welding objects in a GMAW process}
\label{schema-macro-drop-droplet-parameters}
\end{center}
\end{figure}

At figure \ref{figure_1} appears the geometrical elements to be study at the macro drop: the macro drop radius at the base $R_{m}$, the macro drop
  center height $h_{m}$, the macro drop volume $V_{m}$, the penetration below the surface level $p$ and the wetting angles:
   $\Theta_{m(al)}$, $\Theta_{m(ar)}$ (apparent) and $\Theta_{m(rl)}$, $\Theta_{m(rr)}$ (real). With the indices $l$ and $r$ refer to left and
    right angles. The indices: $m$ refer to macro drop, $g$ refer to gaz, $l$ refer to liquid, $s$ refer to solid$/$substrate and $a$ refer to the droplet.

@@

To the droplet or metal transfer drop the geometrical parameters to identify are: The estimate volume $V_{a}$ and the estimate radius $R_{a}$. 
    Finally in order to better understand the droplet dynamic \cite{}, the physical elements to identify are: the fall time $t_{a}$ and the average 
    speed $U_{a}$ (see figure \ref{figure_1}).

@@

As was shown in section \ref{image-acquisition-setup}, the camera recording a shadow projection of the welding process. The projection is
     supposed to be a section of the welding process image, orthogonal to the optical path. However, to correct small angle error in the alignment 
     of the camera, a control image (small chessboard image is use (see figure \ref{photo-calibration}). The real image size are taking from the know width of the welding wire.

\begin{figure}[h!]
\begin{center}    
\subfigure[Weld pool image in static GTAW process]{\label{photo-calibration-cuadro-patron}\includegraphics[width=3.5cm,height=3.5cm]{photo-calibration-cuadro-patron.png}}
\subfigure[Canny filter treatment]{\label{photo-calibration-cuadro-origin}\includegraphics[width=3.5cm,height=3.5cm]{photo-calibration-cuadro-origin.png}}\\
\subfigure[Snake treatment]{\label{photo-calibration-image-origin}\includegraphics[width=3.5cm,height=3.5cm]{photo-explication-snake.png}}
\subfigure[Segmentation image by 2 cluster sample comparison]{\label{photo-explication-segmentation}\includegraphics[width=3.5cm,height=3.5cm]{photo-explication-segmentation.png}}
\end{center}
\caption{{\small Samples of different methods technique for image processing}}
\label{photo-explication}
\end{figure}

@@

@@

To the macro drop profile a user interaction algorithm is responsible to extract the continuous cord (white pixels) corresponding to the profile into a list. Note 
              that the macro drop and the substrate have the same continuous profile (see figure \ref{photo-macro-drop-canny}. The substrate shadow projection 
              is always the same if the optical path doesn’t change. Therefore the users have only to choose, to first image in the experience series, one extreme 
              of the cord corresponding to substrate profile (see figure \ref{photo-macro-drop-contour}). Then the algorithm finds all new white pixels inside a 
              small neighbor around the pixels selected by the user. Each new white pixel
               is added to the profile list. Starting at each new pixel, this process continues, automatically, until the other extreme of the cord, until the last image of the series.

@@

Give $S$ the set of points in the space image. The segments could be building between wherever two points of $S$. However, to fulfill the cords with 
                segments, only the points placed together could define a segment. To choose the correct points of $S$ a Delaunay triangulation is made 
                over $S$ \cite{CGAL} (see figure \ref{photo-droplet-triangulation-delaunay}). Each point of $S$ is a vertices of a triangle and then each potential 
                segment is a side of a triangle. The Delaunay triangulations guarantee a unique solution with minimal internal angles for each triangle, which means 
                minimal length sides to the triangles \cite{GEOMETRICAL*}. Give $\beta$ the cord surrounding the external hull of the Delaunay triangulation of $S$. 
                An alpha shape algorithm uses a $\alpha \in \mathbb{N}$ parameter, or square radius of a test circle, to travel from outside $\beta$ to inside 
                the triangulation. The circle has to go through the triangles sides without touch the vertices. Give $n$ the number of triangles sides of triangulation
                 and $l_{i}$ with $i\in\{1,n\}$ the length of each side, if $\sqrt{\alpha} > l_{i}$ with
                 $i\in\{1,n}$, a segment is build between the respective vertices or points. At the end of the process, a segments set know as $\alpha$-shape has
                  been created.  Note if $\sqrt{alpha} > l_{i} \forall i$, the circle can’t be go inside $\beta$, then $\alpha$-shape is $\beta$ \cite{CGAL-library}. 
                  If $\sqrt{alpha} < l_{i} \forall i$ the circle go through every side in the triangulation, 
                 then no segment are created and $\alpha$-shape became $S$. The points that compound the droplet section are all contact neighbors. Then $\alpha = 1$ 
                 is sufficient to create a segment set around the droplet (see figure \ref{photo-droplet-alpha-shape}). A graph algorithm connects this segment and finds 
                 the longest closer segment. The longest closer segment corresponds to the metal transfer drop profile (see figure \ref{photo-droplet-contour}).

@@

Now it is possible to determine the geometrical parameters shown at figure \ref{shema-macro-drop-droplet-parameters}. In order to this, it has been developed several
                  functions linked with the BAME and erCv libraries. For example, to wetting angle measures, linear regressions or least squares (parabolic, ellipse, circle) 
                  are taken from the processed macro drop extracted profiles. A loop algorithm applies this method to all frames at the experience to compute the wetting 
                  angles. Same procedure is use to estimate the macro drop radius and volume \cite{CHAPUIS}.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%    INSTITUTE OF PHYSICS PUBLISHING                                   %
%                                                                      %
%   `Preparing an article for publication in an Institute of Physics   %
%    Publishing journal using LaTeX'                                   %
%                                                                      %
%    LaTeX source code `ioplau2e.tex' used to generate `author         %
%    guidelines', the documentation explaining and demonstrating use   %
%    of the Institute of Physics Publishing LaTeX preprint files       %
%    `iopart.cls, iopart12.clo and iopart10.clo'.                      %
%                                                                      %
%    `ioplau2e.tex' itself uses LaTeX with `iopart.cls'                %
%                                                                      %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%
% First we have a character check
%
% ! exclamation mark    " double quote  
% # hash                ` opening quote (grave)
% & ampersand           ' closing quote (acute)
% $ dollar              % percent       
% ( open parenthesis    ) close paren.  
% - hyphen              = equals sign
% | vertical bar        ~ tilde         
% @ at sign             _ underscore
% { open curly brace    } close curly   
% [ open square         ] close square bracket
% + plus sign           ; semi-colon    
% * asterisk            : colon
% < open angle bracket  > close angle   
% , comma               . full stop
% ? question mark       / forward slash 
% \ backslash           ^ circumflex
%
% ABCDEFGHIJKLMNOPQRSTUVWXYZ 
% abcdefghijklmnopqrstuvwxyz 
% 1234567890
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
\documentclass[12pt]{iopart}
\usepackage{graphicx}
\usepackage[sf]{subfigure}
\newcommand{\gguide}{{\it Preparing graphics for IOP journals}}
%Uncomment next line if AMS fonts required
%\usepackage{iopams}

\begin{document}

@@

\address{IOP 
Publishing, Dirac
House, Temple Back, Bristol BS1 6BE, UK}
\ead{custserv@iop.org}

\begin{abstract}

@@

\section{Introduction}

A better comprehension of welding objects behavior presents in a PGMAW process
 as macro drop and metal transfer drop (droplet) could help to improve numerical simulations
 and enhance welding quality in manufacturing process \cite{LIN}, \cite{WU1}.
% An example to this are the problems linked to productivity. To some process,
 For example, the
 increase of productivity means often increase of weld speed,
 however some welding defects related to bad shape, as the hump formations, limit the maximum weld 
 speed \cite{CHAPUIS,Mendez} and therefore the productivity. 
 In a PGMAW process, the weld quality is strongly relates with the metal transfer  stability 
 (the droplet deposition). The control of metal transfer process %implies control of
 will improve weld quality \cite{WANG}. In both example cases a macro drop kinetic
  study and a droplet dynamic analysis during static PGMAW process could help to better
 understand some GMAW related phenomena as humping formation \cite{CHO}, droplet 
  fluctuations \cite{WANG} that guarantee the welding quality. 
 In both case, the kinetic and the dynamic analysis, required the monitoring \cite{} and the 
  geometrical analysis of macro drop and droplets in order to extract 
  qualitative information of these welding objects. 
  There have been many studies on visual sensing techniques for observing weld pool 
  image \cite{BAE} and metal transfer process during welding \cite{LIN}. Optical sensors like high speed 
 cameras and lighting systems have been widely use in GMAW process to realize 
 image acquisition \cite{GUANGJUN},
   control process \cite{BAE}, parametric studies \cite{BALSAMO} and droplet dynamics analysis\cite{LIN}.
   Image processing plays a critical role in 
   extracting useful information from visual scenes \cite{WANG}.
 Nevertheless the strong interference from the arc lightning required more than standard 
   image treatment to analyses the raw images of the welding process \cite{NORDBRUCH}. 
Previous work has shown that is possible to perform geometrical 
   analysis in weld pools or droplets \cite{WU2}.
   Parameters such as macro drop or droplets surface, volumes 
or height has been measure using different and 
   specifics processing images algorithms \cite{WANG}, \cite{WU2}, \cite{SEED},
 \cite{NORDBRUCH}. However, to date, effectively automatic images processing
    of metal transfer has not been developed, possibly due to the  difficulty involved 
for welding researchers \cite{WANG}.

@@

\section{Problems description}

Bradstreet was the first researcher to experimentally study the bead hump formation \cite{CHO}. 
Humping was defined as the series of undulations of the weld bead. Further studies as computer 
simulation made by CHO et al. \cite{CHO} shown a relation between the droplet momentum deposition
 and the hump size. Same relations has been found between metal transfer process and welding quality \cite{}

Then the shape and size kinetic analysis of macro drop and droplets could help to better understand 
 and enhance the PGMAW process. To monitoring the shape and size of these welding objects, to a 
 first step, a 2D approach would be sufficient. Therefore a shadowgraphy technique, or back lighting,
  is the natural choice to record the droplets and macro drops profiles \cite{BALSAMO}, furthermore is 
  one of the only way to access these quantities. Due to the arc light interference and the relatively high 
  speed of wire feed process, a high speed camera and an effective image processing algorithms are 
  required \cite{WANG}. The typically frame rate acquisition to metal transfer drop kinetic analysis is $3000$ 
  per second \cite{WANG}. Therefore the algorithms have to be able to extract the geometrical information (area, size and others)
   from the macro drop and droplets, from a huge amount of data. In addition, the voltage and current signals are directly related 
   with the droplet formation at the wire \cite{BALSAMO}. Good synchronizations methods have to be applied between shadowgraphies 
   frames and electrical signals.

\section{Image processing background}

Some definitions and brief description of generals principles used in
 image processing are required to better understand how is work the 
algorithms utilized by the erCv library.

@@

\subsection{Filters}\label{filters}

The filters are algorithms that operate as mathematical functions $f$ above the $X$, $Y$ or both axis of the image ($Z = f(X,Y)$,  
$f(X)$ or $f(Y)$)  modifying his greyness value or $Z$ component. 
Different kind of filters can be mentioned as median, Gaussian, impulse, adaptive and others. The impulse filters such Canny are 
widely use to edge detection \cite{COCQUEREZ}. These filter have an impulse response to most important greyness gradient in the 
image; this allows the filter a better edges localisation (see figure \ref{photo-explication-filter}). For this reason, Canny filter is widely 
use at erCv library to detect the welding elements edges. However, it is sensible to noise or secondary greyness gradients in the image
 and, in consequence, it have some difficulties to define closer surface.

@@

\begin{figure}[h!]
\begin{center}    
\subfigure[Weld pool image in static GTAW process]{\label{photo-explication-patron}\includegraphics[width=3.5cm,height=3.5cm]{photo-explication-patron.png}}
\subfigure[Canny filter treatment]{\label{photo-explication-filter}\includegraphics[width=3.5cm,height=3.5cm]{photo-explication-filter.png}}\\
\subfigure[Snake treatment]{\label{photo-explication-snake}\includegraphics[width=3.5cm,height=3.5cm]{photo-explication-snake.png}}
\subfigure[Segmentation image by 2 cluster sample comparison]{\label{photo-explication-segmentation}\includegraphics[width=3.5cm,height=3.5cm]{photo-explication-segmentation.png}}
\end{center}
\caption{{\small Samples of different methods technique for image processing}}
\label{photo-explication}
\end{figure}

@@

\begin{figure}
\begin{center}
\includegraphics[width=7cm,height=4cm]{schema-erCv.png}
\caption{{\small Flow diagram of erCv library composition}}
\label{schema-erCv}
\end{center}
\end{figure}

@@

\begin{table}
\begin{center}
\begin{tabular}{|cc|}
\hline
Welding wire type & ER70S \\ 
Wire diameter ($mm$) or $drw$ & 1 \\  
Contact tip to work distance ($mm$) & 20 \\
Shielding gaz flowrate ($l/min$) & 18 \\ \hline
\end{tabular}
\caption{{\small Variable welding parameters used in experiments}}
\label{table-parameters-change}
\end{center}
\end{table}

@@

At figure \ref{schema-macro-drop-droplet-parameters} appears the geometrical elements to be study at the macro drop: the macro drop radius at the base $R_{m}$, the macro drop
  center height $h_{m}$, the macro drop volume $V_{m}$, the penetration below the surface level $p$ and the wetting angles:
   $\Theta_{m(al)}$, $\Theta_{m(ar)}$ (apparent) and $\Theta_{m(rl)}$, $\Theta_{m(rr)}$ (real). With the indices $l$ and $r$ refer to left and
    right angles. The indices: $m$ refer to macro drop, $g$ refer to gaz, $l$ refer to liquid, $s$ refer to solid$/$substrate and $a$ refer to the droplet.

\begin{figure}
\begin{center}
\includegraphics[width=7cm,height=4cm]{schema-macro-drop-droplet-parameters.png}
\caption{{\small Profile schematics of welding objects in a GMAW process}}
\label{schema-macro-drop-droplet-parameters}
\end{center}
\end{figure}

@@

To the droplet or metal transfer drop the geometrical parameters to identify are: The estimate volume $V_{a}$ and the estimate radius $R_{a}$. 
    Finally in order to better understand the droplet dynamic \cite{}, the physical elements to identify are: the fall time $t_{a}$ and the average 
    speed $U_{a}$ (see figure \ref{schema-macro-drop-droplet-parameters}).

@@

As was shown in section \ref{image-acquisition-setup}, the camera recording a shadow projection of the welding process. The projection is
     supposed to be a section of the welding process image, orthogonal to the optical path. Therefore, to a first approach, not image projection 
     correction is necessary. Only to calibrate the image dimension (at pixels) to the real objects dimension (at millimeters), a conversion scale is 
     made with the known width of the welding wire $conv = 1\ mm/ 20\ pixels$ .

@@

\begin{figure}[h!]
\begin{center}    
\subfigure[Macro drop raw image]{\label{photo-macro-drop-patron}\includegraphics[width=3.5cm,height=3.5cm]{photo-macro-drop-original.png}}
\subfigure[Macro drop grayness image]{\label{photo-macro-drop-grey}\includegraphics[width=3.5cm,height=3.5cm]{photo-macro-drop-grey.png}}\\
\subfigure[First smooth filter]{\label{photo-macro-drop-smoothblur}\includegraphics[width=3.5cm,height=3.5cm]{photo-macro-drop-smoothblur.png}}
\subfigure[Adaptive threshold per regions]{\label{photo-macro-drop-adaptivethreshold}\includegraphics[width=3.5cm,height=3.5cm]{photo-macro-drop-adaptivethreshold.png}}\\
\subfigure[Median filter]{\label{photo-macro-drop-smoothmedian}\includegraphics[width=3.5cm,height=3.5cm]{photo-macro-drop-smoothmedian.png}}
\subfigure[Canny filter]{\label{photo-macro-drop-canny}\includegraphics[width=3.5cm,height=3.5cm]{photo-macro-drop-canny.png}}\\
\end{center}
\caption{{\small Image treatment step by step of PGMAW process}}
\label{photo-macro-drop}
\end{figure}

@@

To the macro drop profile a user interaction algorithm is responsible to extract the continuous cord (white pixels) corresponding to the profile into a list. Note 
              that the macro drop and the substrate have the same continuous profile (see figure \ref{photo-macro-drop-canny}. The substrate shadow projection 
              is always the same if the optical path doesn’t change. Therefore the users have only to choose, to first image in the experience series, one extreme 
              of the cord corresponding to substrate profile (see figures \ref{photo-results-macro-drop}). Then the algorithm finds all new white pixels inside a 
              small neighbor around the pixels selected by the user. Each new white pixel
               is added to the profile list. Starting at each new pixel, this process continues, automatically, until the other extreme of the cord, until the last image of the series.

\begin{figure}[h!]
\begin{center}    
\subfigure[t = ]{\label{photo-results-macro-drop-t1}\includegraphics[width=3.5cm,height=3.5cm]{photo-results-macro-drop-t1.png}}
\subfigure[t = ]{\label{photo-results-macro-drop-t2}\includegraphics[width=3.5cm,height=3.5cm]{photo-results-macro-drop-t2.png}}\\
\subfigure[t = ]{\label{photo-results-macro-drop-t3}\includegraphics[width=3.5cm,height=3.5cm]{photo-results-macro-drop-t3.png}}
\subfigure[t = ]{\label{photo-results-macro-drop-t4}\includegraphics[width=3.5cm,height=3.5cm]{photo-results-macro-drop-t4.png}}\\
\end{center}
\caption{{\small Images series of macro drop profiles}}
\label{photo-results-macro-drop}
\end{figure}

@@

Give $S$ the set of points in the space image. The segments could be building between wherever two points of $S$. However, to fulfill the cords with 
                segments, only the points placed together could define a segment. To choose the correct points of $S$ a Delaunay triangulation is made 
                over $S$ \cite{CGAL}. Each point of $S$ is a vertices of a triangle and then each potential 
                segment is a side of a triangle. The Delaunay triangulations guarantee a unique solution with minimal internal angles for each triangle, which means 
                minimal length sides to the triangles \cite{GEOMETRICAL}. Give $\beta$ the cord surrounding the external hull of the Delaunay triangulation of $S$. 
                An alpha shape algorithm uses a $\alpha \in N$ parameter, or square radius of a test circle, to travel from outside $\beta$ to inside 
                the triangulation. The circle has to go through the triangles sides without touch the vertices. Give $n$ the number of triangles sides of triangulation
                 and $l_{i}$ with $i\in\{1,n\}$ the length of each side, if $\sqrt{\alpha} > l_{i}$ with
                 $i \in\{1,n\}$, a segment is build between the respective vertices or points. At the end of the process, a segments set know as $\alpha$-shape has
                  been created.  Note if $\sqrt{\alpha} > l_{i} \forall i$, the circle can’t be go inside $\beta$, then $\alpha$-shape is $\beta$ \cite{CGAL-library}. 
                  If $\sqrt{\alpha} < l_{i} \forall i$ the circle go through every side in the triangulation, 
                 then no segment are created and $\alpha$-shape became $S$. The points that compound the droplet section are all contact neighbors. Then $\alpha = 1$ 
                 is sufficient to create a segment set around the droplet (see figure \ref{photo-droplet-alpha-shape}).

\begin{figure}[h!]
\begin{center}    
\subfigure[t = ]{\label{photo-droplet-canny}\includegraphics[width=3.5cm,height=3.5cm]{photo-droplet-canny.png}}
\subfigure[t = ]{\label{photo-droplet-alpha-shape}\includegraphics[width=3.5cm,height=3.5cm]{photo-droplet-alpha-shape.png}}\\
\end{center}
\caption{{\small Images series of macro drop profiles}}
\label{photo-results-droplet}
\end{figure}

A graph algorithm connects this segment and finds the longest closer segment. The longest closer segment corresponds to the metal transfer drop profile 
(see figure \ref{photo-results-droplet}).

\begin{figure}[h!]
\begin{center}    
\subfigure[t = ]{\label{photo-results-droplet-t1}\includegraphics[width=3.5cm,height=3.5cm]{photo-results-droplet-t1.png}}
\subfigure[t = ]{\label{photo-results-droplet-t2}\includegraphics[width=3.5cm,height=3.5cm]{photo-results-droplet-t2.png}}\\
\subfigure[t = ]{\label{photo-results-droplet-t3}\includegraphics[width=3.5cm,height=3.5cm]{photo-results-droplet-t3.png}}
\subfigure[t = ]{\label{photo-results-droplet-t4}\includegraphics[width=3.5cm,height=3.5cm]{photo-results-droplet-t4.png}}\\
\end{center}
\caption{{\small Images series of macro drop profiles}}
\label{photo-results-droplet}
\end{figure}

@@

Now it is possible to determine the geometrical parameters shown at figure \ref{schema-macro-drop-droplet-parameters}. In order to this, it has been developed several
                  functions linked with the BAME and erCv libraries. For example, to wetting angle measures, linear regressions or least squares (parabolic, ellipse, circle) 
                  are taken from the processed macro drop extracted profiles. A loop algorithm applies this method to all frames at the experience to compute the wetting 
                  angles. Same procedure is use to estimate the macro drop radius and volume \cite{CHAPUIS}.
